---
title: "Plot_table_Patients_Finder_Tool_Report"
author: "Phuong Anh Mai"
date: "2024-10-29"
output: html_document
---

This R markdown files contain all plot and table code for the Patients Finder Tool report. 

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(readr)
library(dplyr)
library(tidymodels)
library(stringdist)
library(glmnet)
library(randomForest)
library(gbm)
library(yardstick)
library(ggplot2)
library(plotly)
library(gridExtra)
library(kableExtra)
library(knitr)
library(visdat)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
```

```{r reading datasets, echo = FALSE, warning=FALSE, message = FALSE}
waiting_list <- read_csv("data/new_waiting_list.csv")
surgery_patients <- read.csv("data/joined_data.csv")
```

# Data wrangling 
## Convert Data types and Filter based on assumptions

```{r wrangling, echo = FALSE, warning=FALSE, message=FALSE}
# Filter out patient > 17 years old
surgery_patients <- surgery_patients %>% 
  filter(AgeAtDateOfRemoval >17)

# Getting waiting list data
waiting_patients <- waiting_list %>% 
  filter(ReportedRemovalDateTime == "NULL",
         RemovalDateTime == "NULL")

# Remove or impute missing data 
surgery_patients <- na.omit(surgery_patients)

# Remove unwanted columns
unwanted <- c("Paed ENT",
"Paed General Surgery",
"Paed Oral & Maxillofacial Surgery",
"Paed Orthopaedic Surgery",
"Paed Plastic Surgery",
"Paed Urology")

surgery_patients <- surgery_patients[!surgery_patients$WaitingListSpecialtyDesc %in% unwanted, ]

# Identify duplicates
duplicates <- duplicated(surgery_patients)
# Remove duplicates
surgery_patients <- surgery_patients[!duplicated(surgery_patients), ]

# glimpse(surgery_patients)

# Create PPPCategory for both datasets
surgery_patients <- surgery_patients %>%
  mutate(PPPCategory = case_when(
    grepl("arthro|joint|shoulder|knee|ligament|rotator cuff|ACL|hip|meniscus|spine|fusion|discectomy", 
          PPPDesc, ignore.case = TRUE) ~ "Orthopedic",
    grepl("heart|cardiac|bypass|coronary|valve|CABG|aortic|angioplasty|stent|aneurysm", 
          PPPDesc, ignore.case = TRUE) ~ "Cardiac",
    grepl("appendic|cholecyst|bowel|hernia|gastrectomy|colectomy|colorectal|laparotomy|gastrostomy|fundoplication", 
          PPPDesc, ignore.case = TRUE) ~ "General Surgery",
    grepl("breast|mastectomy|lumpectomy|reconstruction|augmentation|reduction|prosthesis|biopsy|oncology", 
          PPPDesc, ignore.case = TRUE) ~ "Breast Surgery",
    grepl("urology|kidney|nephrectomy|bladder|prostate|ureter|urethra|renal", 
          PPPDesc, ignore.case = TRUE) ~ "Urology",
    grepl("ENT|sinus|tonsil|adenoid|larynx|otoplasty|nose|nasal|septoplasty|ear|throat", 
          PPPDesc, ignore.case = TRUE) ~ "ENT Surgery",
    grepl("gynaecology|hysterectomy|cervix|uterus|fallopian|ovary|endometrial|ablation|vaginal", 
          PPPDesc, ignore.case = TRUE) ~ "Gynaecology",
    grepl("neurosurgery|brain|craniotomy|spinal|meninges|nerve|tumor|neuro", 
          PPPDesc, ignore.case = TRUE) ~ "Neurosurgery",
    TRUE ~ "Other"  # Default category if none of the above matches
  ))


waiting_list <- waiting_list %>%
  mutate(PPPCategory = case_when(
    grepl("arthro|joint|shoulder|knee|ligament|rotator cuff|ACL|hip|meniscus|spine|fusion|discectomy", 
          PPPDesc, ignore.case = TRUE) ~ "Orthopedic",
    grepl("heart|cardiac|bypass|coronary|valve|CABG|aortic|angioplasty|stent|aneurysm", 
          PPPDesc, ignore.case = TRUE) ~ "Cardiac",
    grepl("appendic|cholecyst|bowel|hernia|gastrectomy|colectomy|colorectal|laparotomy|gastrostomy|fundoplication", 
          PPPDesc, ignore.case = TRUE) ~ "General Surgery",
    grepl("breast|mastectomy|lumpectomy|reconstruction|augmentation|reduction|prosthesis|biopsy|oncology", 
          PPPDesc, ignore.case = TRUE) ~ "Breast Surgery",
    grepl("urology|kidney|nephrectomy|bladder|prostate|ureter|urethra|renal", 
          PPPDesc, ignore.case = TRUE) ~ "Urology",
    grepl("ENT|sinus|tonsil|adenoid|larynx|otoplasty|nose|nasal|septoplasty|ear|throat", 
          PPPDesc, ignore.case = TRUE) ~ "ENT Surgery",
    grepl("gynaecology|hysterectomy|cervix|uterus|fallopian|ovary|endometrial|ablation|vaginal", 
          PPPDesc, ignore.case = TRUE) ~ "Gynaecology",
    grepl("neurosurgery|brain|craniotomy|spinal|meninges|nerve|tumor|neuro", 
          PPPDesc, ignore.case = TRUE) ~ "Neurosurgery",
    TRUE ~ "Other"  # Default category 
  ))

# Convert data types
surgery_patients <- surgery_patients %>%
  mutate(
    PlannedTheatreTimeInMins = as.numeric(PlannedTheatreTimeInMins),
    ArrivedTheatreDateTime = as.POSIXct(ArrivedTheatreDateTime, format = "%Y-%m-%d %H:%M:%S"),
    OutOfTheatreDateTime = as.POSIXct(OutOfTheatreDateTime, format = "%Y-%m-%d %H:%M:%S"),
    WaitingStartDateTime = as.POSIXct(WaitingStartDateTime, format = "%Y-%m-%d %H:%M:%S"),
    SurgeryStartDateTime = as.POSIXct(SurgeryStartDateTime, format = "%Y-%m-%d %H:%M:%S"),
    SurgeryEndDateTime = as.POSIXct(SurgeryEndDateTime, format = "%Y-%m-%d %H:%M:%S"),
    IntoTheatreDateTime = as.POSIXct(IntoTheatreDateTime, format = "%Y-%m-%d %H:%M:%S"),
    AgeGroupAtDateOfRemoval = as.factor(AgeGroupAtDateOfRemoval),
    CurrentPriorityCode = as.factor(CurrentPriorityCode),
    PPPCategory = as.factor(PPPCategory),
    # Use for model fitting
    WaitingListSpecialtyDesc = as.factor(WaitingListSpecialtyDesc),
    OperatingDrCarerName = as.factor(OperatingDrCarerName),
    TreatmentCampusName = as.factor(TreatmentCampusName),
    PlannedStayTypeDescription = as.factor(PlannedStayTypeDescription),
    AdminCategory = as.factor(AdminCategory),
  )

# Calculate real-time of surgery
surgery_patients <- surgery_patients %>% 
  mutate(SurgeryLength = round((difftime(OutOfTheatreDateTime, IntoTheatreDateTime, units = "mins")),4)) %>% 
  mutate(SurgeryLength = as.numeric((SurgeryLength)))

# Convert Surgery Length data types
surgery_patients <- surgery_patients %>%
mutate(SurgeryLength = as.numeric(SurgeryLength))

waiting_list <- waiting_list %>%
  mutate(
    AgeGroupAtDateOfRemoval = as.factor(AgeGroupAtDateOfRemoval),
    CurrentPriorityCode = as.factor(CurrentPriorityCode),
    WaitingListSpecialtyDesc = as.factor(WaitingListSpecialtyDesc),
    PPPCategory = as.factor(PPPCategory),
    TreatmentCampusName = as.factor(TreatmentCampusName),
    PlannedStayTypeDescription = as.factor(PlannedStayTypeDescription),
    AdminCategory = as.factor(AdminCategory)
  )
```

## Further refine of the dataset

```{r surgery length, echo = FALSE, warning=FALSE, message=FALSE}}
# Remove Specialty with under 10 cases
unwanted_specialty<- surgery_patients %>% 
  group_by(WaitingListSpecialtyDesc) %>%
  summarise(CaseCount = n()) %>% 
  filter(CaseCount < 10)

surgery_patients <- surgery_patients[!surgery_patients$WaitingListSpecialtyDesc %in% unwanted_specialty$WaitingListSpecialtyDesc, ]

# Remove negative values and outliers
# Filter out Surgery within 10 hours, as the longer surgery requested more complex estimation
surgery_patients <- surgery_patients %>% 
  na.omit() %>% 
  filter(SurgeryLength > 10) %>% 
  filter(SurgeryLength < 600) %>% 
  filter(PlannedTheatreTimeInMins > 10) %>% 
  filter(PlannedTheatreTimeInMins < 600)
```

# Exploratory Data Analysis (EDA)
## Checking the completeness of the data after clean

```{r data completeness, echo = FALSE, warning=FALSE, message=FALSE}}
p1 <- vis_dat(surgery_patients, `warn_large_data` = FALSE)

p2 <- vis_dat(waiting_patients, `warn_large_data` = FALSE)

plot_grid(p1, p2) 
```

## Observe the human estimation and real-time of Surgery Length

```{r distribution visualize, echo = FALSE, warning=FALSE, message=FALSE}
# Visualize the distribution of Planned Theatre Time
a <- ggplot(surgery_patients, aes(x = PlannedTheatreTimeInMins)) +
  geom_histogram(fill = "blue", color = "white", bins = 30) +
  theme_minimal() +
  labs(title = "Distribution of Planned Theatre Time in Minutes",
       x = "Planned Theatre Time (mins)")

# Visualize the distribution of Time Difference
b <- ggplot(surgery_patients, aes(x = SurgeryLength)) +
  geom_histogram( fill = "#E69F00", color = "white", bins = 30) +
  theme_minimal() +
  labs(title = "Distribution of Surgery Time in Minutes",
       x = "Operating Theatre Time (mins)")

gridExtra::grid.arrange(a, b)
```

## Investigate the Surgery Length by Specialty

```{r visualize boxplot, echo = FALSE, warning=FALSE, message=FALSE}
# Boxplot of SurgeryLength with an additional geom_line for average
ggplot(surgery_patients, aes(x = SurgeryLength, y = fct_reorder(WaitingListSpecialtyDesc, SurgeryLength, .fun = median))) +
  geom_boxplot(fill = "#E69F00") +
  
  theme_minimal() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 55, hjust = 1)) +
  labs(title = "Surgery Length by Specialty",
       x = "Surgery Length (mins)",
       y = "Specialty")
```

## Observe the Waiting Days

```{r Waiting days, echo = FALSE, warning=FALSE, message=FALSE}
# Calculating Waiting day variables 
surgery_patients <- surgery_patients %>%
  mutate(WaitingDays = as.numeric(round(difftime(SurgeryStartDateTime, WaitingStartDateTime, units = "days"))))

# Proportion of people in waiting within 1 year
proportion_within_1_year <- surgery_patients %>%
  summarise(Proportion = round(mean(WaitingDays <= 365),4))

# Average waiting time for particular Specialty
Average_waiting_day_by_specialty <- surgery_patients %>%
  group_by(WaitingListSpecialtyDesc) %>%
  summarise(Avearge_waiting_day = round(mean(WaitingDays)),
            CaseCount = n()) %>% 
  arrange(desc(Avearge_waiting_day))

# Plot Average waiting time for particular Specialty
p <- ggplot(Average_waiting_day_by_specialty, aes(x = reorder(WaitingListSpecialtyDesc, Avearge_waiting_day), y = Avearge_waiting_day)) +
  geom_segment(aes(xend = reorder(WaitingListSpecialtyDesc, Avearge_waiting_day), y = 0, yend = Avearge_waiting_day), color = "grey") +
  geom_point(aes(size = CaseCount), color = "#E69F00")  +
  theme_minimal() +
  labs(title = "Average Waiting Day by Specialty",
       x = "Specialty",
       y = "Avearge waiting day") +
  theme(axis.text.x = element_text(angle = 55, hjust = 1))

plotly_plot <- ggplotly(p)

plotly_plot
```


# Feature engineering: Extract interesting features
## Calculate and plot the Average Surgery length by Age Group

```{r}
# Calculating Average Surgery Length by Age Group
AverageSurgeryLength_age <- surgery_patients %>% 
  group_by(AgeGroupAtDateOfRemoval) %>% 
  summarise(AverageSurgeryLength_age = round(mean(SurgeryLength, na.rm = TRUE))) %>% 
  arrange(desc(AverageSurgeryLength_age))

# Displaying the result
ggplot(AverageSurgeryLength_age, aes(x = AgeGroupAtDateOfRemoval, y = AverageSurgeryLength_age)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength_age), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Age Group", 
       x = "Age Groups", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10)
  )
```

## Calculate and plot the Average Surgery length by Specialty

```{r}
# Calculate the average of SurgeryLength by specialty
Average_surgery_length_by_specialty <- surgery_patients %>%
  group_by(WaitingListSpecialtyDesc) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the result
ggplot(Average_surgery_length_by_specialty, aes(x = reorder(WaitingListSpecialtyDesc, -AverageSurgeryLength), y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Specialty", 
       x = "Specialty", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10)
  )
```

## Calculate and plot the Average Surgery length by Admin category

```{r}
# Calculate average surgery length by Admin category
Average_surgery_length_admin <- surgery_patients %>%
  group_by(AdminCategory) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the result
ggplot(Average_surgery_length_admin, aes(x = reorder(AdminCategory, -AverageSurgeryLength), y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Admin Category", 
       x = "Admin Category", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10) 
  )
```

## Calculate and plot the Average Surgery length by Planned Stay

```{r}
# Calculate average surgery length by Planned Stay
Average_surgery_length_plan_stay <- surgery_patients %>%
  group_by(PlannedStayTypeDescription) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Displaying the result
Average_surgery_length_plan_stay %>%
  kable("html", caption = "Average Surgery Length by Planned Stay") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = FALSE, 
                font_size = 14) %>%
  column_spec(1, bold = TRUE, border_right = TRUE) %>%  
  column_spec(2, color = "blue", background = "lightgrey") %>% 
  add_header_above(c(" " = 1, " " = 1)) 
```

## Calculate and plot the Average Surgery length by Treatment Campus

```{r}
# Calculate average surgery length by Treatment Campus
Average_surgery_length_treatment_campus <- surgery_patients %>%
  group_by(TreatmentCampusName) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the results
ggplot(Average_surgery_length_treatment_campus, aes(x = reorder(TreatmentCampusName, -AverageSurgeryLength), y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Treatment Campus", 
       x = "Treatment Campus", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10)
  )
```

## Calculate and plot the Average Surgery length by Current Priority Codes

```{r}
# Calculate average surgery length by Current Priority Codes
Average_surgery_length_priority <- surgery_patients %>%
  group_by(CurrentPriorityCode) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the results
ggplot(Average_surgery_length_priority, aes(x = CurrentPriorityCode, y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Priority Code", 
       x = "Priority Code", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10)
  )
```

## Calculate and plot the Average Surgery Length by PPP Category

```{r PPP Desc by group}
# Calculate average surgery length by PPP Category
Average_surgery_length_ppp_category <- surgery_patients %>%
  group_by(PPPCategory) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the result
ggplot(Average_surgery_length_ppp_category, aes(x = reorder(PPPCategory, -AverageSurgeryLength), y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +  
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +  
  coord_flip() +  
  labs(title = "Average Surgery Length by PPP Category", 
       x = "PPP Category", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +  # Clean theme
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.text.y = element_text(size = 12),  
    axis.title.y = element_text(size = 14),  
    axis.title.x = element_text(size = 14),  
    plot.margin = ggplot2::margin(10, 10, 10, 10)  
  )
```

# Fitting Model
## Split data

```{r setting up, echo = FALSE, warning=FALSE, message=FALSE}
set.seed(123)
# Training and Testing sets
 surgery_patients_split <- initial_split(surgery_patients,
                                        2/3,
                                        strata = WaitingListSpecialtyDesc)

 train_data <- training(surgery_patients_split)
 test_data <- testing(surgery_patients_split)
    
# Set consistent factor levels in train and test data
age_levels <- levels(train_data$AgeGroupAtDateOfRemoval)
priority_levels <- levels(train_data$CurrentPriorityCode)
specialty_levels <- levels(train_data$WaitingListSpecialtyDesc)
ppp_category <- levels(train_data$PPPCategory)
treament_campus_levels <- levels(train_data$TreatmentCampusName)
planned_stay_levels <- levels(train_data$PlannedStayTypeDescription)
admin_levels <- levels(train_data$AdminCategory)

test_data <- test_data %>%
  mutate(
    AgeGroupAtDateOfRemoval = factor(AgeGroupAtDateOfRemoval, levels = age_levels),
    CurrentPriorityCode = factor(CurrentPriorityCode, levels = priority_levels),
    WaitingListSpecialtyDesc = factor(WaitingListSpecialtyDesc, levels = specialty_levels),
    PPPCategory = factor(PPPCategory, levels = ppp_category),
    TreatmentCampusName = factor(TreatmentCampusName, levels = treament_campus_levels),
    PlannedStayTypeDescription = factor(PlannedStayTypeDescription, levels = planned_stay_levels),
    AdminCategory = factor(AdminCategory, levels = admin_levels)
    )

```

## Using cv to find the best hyperparameter for each model

```{r cross_validation, echo = FALSE, warning=FALSE, message=FALSE}}
set.seed(123)
# Create cross-validation folds 
surgery_folds <- vfold_cv(train_data, v = 5, strata = WaitingListSpecialtyDesc)

# Define hyperparameter grids(Tuning)
# Random Forest hyperparameter grid
rf_grid <- expand.grid(
  mtry = c(2, 3, 4, 5),
  min_n = c(5, 10, 15),
  ntree = c(500)
)

# GBM hyperparameter grid
gbm_grid <- expand.grid(
  n.trees = c(100, 300, 500),
  interaction.depth = c(2, 3, 4),
  shrinkage = c(0.01, 0.05, 0.1),
  n.minobsinnode = c(5, 10, 15)
)

# Lambda grid for Ridge and Lasso regression
lambda_grid <- 10^seq(-4, 1, length = 100)

# Define performance metrics
# metrics <- metric_set(rmse, mae, mape, rsq)
metrics <- metric_set(rmse, mae, rsq)

# Train models with hyperparameter tuning on training set of each fold
surgery_models <- surgery_folds %>% 
  mutate(
    # Prepare data matrices for glmnet
    data_matrices = map(splits, ~{
      train_data <- analysis(.x)
      x_train <- model.matrix(log(SurgeryLength) ~ AgeGroupAtDateOfRemoval + CurrentPriorityCode +
                                WaitingListSpecialtyDesc + 
                                PlannedStayTypeDescription +
                                PPPCategory +
                                TreatmentCampusName + AdminCategory, data = train_data)[, -1]
      y_train <- log(train_data$SurgeryLength)
      list(x_train = x_train, y_train = y_train)
    }),
    
    # Linear Regression Model
    reg = map(splits, ~lm(log(SurgeryLength)~ AgeGroupAtDateOfRemoval + CurrentPriorityCode +
                                WaitingListSpecialtyDesc + 
                                PlannedStayTypeDescription +
                                PPPCategory +
                                TreatmentCampusName + AdminCategory,
                          data = analysis(.x))),
    
    
    # Ridge Regression with Hyperparameter Tuning
    ridge = map(data_matrices, ~{
      cv_ridge <- cv.glmnet(
        x = .x$x_train,
        y = .x$y_train,
        alpha = 0,
        lambda = lambda_grid,
        nfolds = 5
      )
      # Return the model with best lambda
      list(model = cv_ridge, best_lambda = cv_ridge$lambda.min)
    }),
    
    # Lasso Regression with Hyperparameter Tuning
    lasso = map(data_matrices, ~{
      cv_lasso <- cv.glmnet(
        x = .x$x_train,
        y = .x$y_train,
        alpha = 1,
        lambda = lambda_grid,
        nfolds = 5
      )
      # Return the model with best lambda
      list(model = cv_lasso, best_lambda = cv_lasso$lambda.min)
    }),
    
    # Random Forest Model with Hyperparameter Tuning
    rf = map(splits, ~{
      train_data <- analysis(.x)
      test_data <- assessment(.x)
            
      
      # Grid search for Random Forest
      rf_results <- rf_grid %>%
        mutate(
          model = pmap(list(mtry, min_n, ntree), function(mtry, min_n, ntree) {
            randomForest(
              log(SurgeryLength) ~ AgeGroupAtDateOfRemoval + CurrentPriorityCode +
                                WaitingListSpecialtyDesc + 
                                PlannedStayTypeDescription +
                                PPPCategory +
                                TreatmentCampusName + AdminCategory,
              data = train_data,
              mtry = mtry,
              nodesize = min_n,
              ntree = ntree,
              importance = TRUE
            )
          }),
          # Evaluate on validation set using yardstick::rmse()
          rmse = map_dbl(model, ~{
            pred_log <- predict(.x, newdata = test_data)
            pred <- exp(pred_log)
            actual <- test_data$SurgeryLength
            df <- data.frame(actual = actual, pred = pred)
            yardstick::rmse(df, truth = actual, estimate = pred)$.estimate
          })
        )
      
      # Select the model with the lowest RMSE
      best_rf <- rf_results %>%
        arrange(rmse) %>%
        slice(1) %>%
        pull(model) %>%
        .[[1]]
      
      # Store best hyperparameters
      best_params <- rf_results %>%
        arrange(rmse) %>%
        slice(1) %>%
        select(mtry, min_n, ntree)
      
      list(model = best_rf, best_params = best_params)
    }),
    
    # GBM Model with Hyperparameter Tuning
    gbm = map(splits, ~{
      train_data <- analysis(.x)
      test_data <- assessment(.x)
      
      # Grid search for GBM
      gbm_results <- gbm_grid %>%
        mutate(
          model = pmap(list(n.trees, interaction.depth, shrinkage, n.minobsinnode), function(n.trees, interaction.depth, shrinkage, n.minobsinnode) {
            gbm(
              formula = log(SurgeryLength) ~ AgeGroupAtDateOfRemoval + CurrentPriorityCode +
                                WaitingListSpecialtyDesc + 
                                PlannedStayTypeDescription +
                                PPPCategory +
                                TreatmentCampusName + AdminCategory,
              data = train_data,
              distribution = "gaussian",
              n.trees = n.trees,
              interaction.depth = interaction.depth,
              shrinkage = shrinkage,
              n.minobsinnode = n.minobsinnode,
              verbose = FALSE
            )
          }),
          # Evaluate on validation set using yardstick::rmse()
          rmse = map_dbl(model, ~{
            pred_log <- predict(.x, newdata = test_data, n.trees = .x$n.trees)
            pred <- exp(pred_log)
            actual <- test_data$SurgeryLength
            df <- data.frame(actual = actual, pred = pred)
            yardstick::rmse(df, truth = actual, estimate = pred)$.estimate
          })
        )
      
      # Select the model with the lowest RMSE
      best_gbm <- gbm_results %>%
        arrange(rmse) %>%
        slice(1) %>%
        pull(model) %>%
        .[[1]]
      
      # Store best hyperparameters
      best_params <- gbm_results %>%
        arrange(rmse) %>%
        slice(1) %>%
        select(n.trees, interaction.depth, shrinkage, n.minobsinnode)
      
      list(model = best_gbm, best_params = best_params)
    })
  )
```

## Prediction for cv test 

```{r cv_pred, echo = FALSE, warning=FALSE, message=FALSE}
# Make predictions on the testing set of cv
surgery_predictions <- surgery_models %>%
  mutate(
    # Prepare testing data matrices
    test_matrices = map(splits, ~{
      test_data <- assessment(.x)
      x_test <- model.matrix(
        log(SurgeryLength)~ AgeGroupAtDateOfRemoval + CurrentPriorityCode +
                                WaitingListSpecialtyDesc + 
                                PlannedStayTypeDescription +
                                PPPCategory +
                                TreatmentCampusName + AdminCategory, data = test_data)[, -1]
      y_test <- test_data$SurgeryLength
      list(x_test = x_test, y_test = y_test)
    }),
    
     # Predictions for Linear Regression
    reg_pred = map2(reg, splits, ~{
      test_data <- assessment(.y)
      pred_log <- predict(.x, newdata = test_data)
      pred <- exp(pred_log)
      actual <- test_data$SurgeryLength
      data.frame(actual = actual, predicted = pred)
    }),
    
    # Predictions for Ridge Regression
    ridge_pred = map2(ridge, test_matrices, ~{
      pred_log <- predict(.x$model, newx = .y$x_test, s = .x$best_lambda)
      pred <- exp(pred_log)
      actual <- .y$y_test
      data.frame(actual = actual, predicted = as.vector(pred))
    }),
    
    # Predictions for Lasso Regression
    lasso_pred = map2(lasso, test_matrices, ~{
      pred_log <- predict(.x$model, newx = .y$x_test, s = .x$best_lambda)
      pred <- exp(pred_log)
      actual <- .y$y_test
      data.frame(actual = actual, predicted = as.vector(pred))
    }),
    
    # Predictions for Random Forest
    rf_pred = map2(rf, splits, ~{
      test_data <- assessment(.y)
      pred_log <- predict(.x$model, newdata = test_data)
      pred <- exp(pred_log)
      actual <- test_data$SurgeryLength
      data.frame(actual = actual, predicted = pred)
    }),
    
    # Predictions for GBM
    gbm_pred = map2(gbm, splits, ~{
      test_data <- assessment(.y)
      pred_log <- predict(.x$model, newdata = test_data, n.trees = .x$model$n.trees)
      pred <- exp(pred_log)
      actual <- test_data$SurgeryLength
      data.frame(actual = actual, predicted = pred)
    })
  )
```

```{r cv_metrics, echo = FALSE, warning=FALSE, message=FALSE}
# Calculate metrics
surgery_metrics <- surgery_predictions %>%
  mutate(
    # Metrics for Linear Regression
    reg_metrics = map(reg_pred, ~{
      yardstick::metrics(.x, truth = actual, estimate = predicted)
    }),
    
    # Metrics for Ridge Regression
    ridge_metrics = map(ridge_pred, ~{
      yardstick::metrics(.x, truth = actual, estimate = predicted)
    }),
    
    # Metrics for Lasso Regression
    lasso_metrics = map(lasso_pred, ~{
      yardstick::metrics(.x, truth = actual, estimate = predicted)
    }),
    
    # Metrics for Random Forest
    rf_metrics = map(rf_pred, ~{
      yardstick::metrics(.x, truth = actual, estimate = predicted)
    }),
    
    # Metrics for GBM
    gbm_metrics = map(gbm_pred, ~{
      yardstick::metrics(.x, truth = actual, estimate = predicted)
    })
  )
```

```{r metrics_combine, echo = FALSE, warning=FALSE, message=FALSE}
# Combine metrics and reshape data
reg_metrics_wide <- surgery_metrics %>% 
  select(id, reg_metrics) %>%
  unnest(cols = reg_metrics) %>%
  select(id, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(model = 'reg')

ridge_metrics_wide <- surgery_metrics %>% 
  select(id, ridge_metrics) %>%
  unnest(cols = ridge_metrics) %>%
  select(id, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(model = 'ridge')

lasso_metrics_wide <- surgery_metrics %>% 
  select(id, lasso_metrics) %>%
  unnest(cols = lasso_metrics) %>%
  select(id, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(model = 'lasso')

rf_metrics_wide <- surgery_metrics %>% 
  select(id, rf_metrics) %>%
  unnest(cols = rf_metrics) %>%
  select(id, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(model = 'rf')

gbm_metrics_wide <- surgery_metrics %>% 
  select(id, gbm_metrics) %>%
  unnest(cols = gbm_metrics) %>%
  select(id, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(model = 'gbm')

surgery_metrics_wide <- bind_rows(
  reg_metrics_wide,
  ridge_metrics_wide,
  lasso_metrics_wide,
  rf_metrics_wide,
  gbm_metrics_wide
)

surgery_results <- surgery_metrics_wide %>% 
  group_by(model) %>% 
  summarize(
    rmse = mean(rmse, na.rm = TRUE),
    mae = mean(mae, na.rm = TRUE),
    rsq = mean(rsq, na.rm = TRUE)
  )

print(surgery_results)

```

```{r Visualize metrics, echo = FALSE, warning=FALSE, message=FALSE}
# Visualize metrics across models using parallel coordinate plots
surgery_metrics_wide %>%
  select(id, model, rmse, mae, rsq) %>%
  GGally::ggparcoord(
    columns = 3:5,  
    groupColumn = "model",  
    showPoints = TRUE,
    scale = "globalminmax"
  ) +
  labs(x = "Metric", y = "Value", title = "Model Performance Across Metrics") +
  theme_minimal()
```

```{r best_hyperparameters, echo = FALSE, warning=FALSE, message=FALSE}
# Extract best hyperparameters for Ridge Regression
ridge_best_lambdas <- surgery_models %>%
  mutate(best_lambda = map_dbl(ridge, "best_lambda")) %>%
  select(id, best_lambda)

# Extract best hyperparameters for Lasso Regression
lasso_best_lambdas <- surgery_models %>%
  mutate(best_lambda = map_dbl(lasso, "best_lambda")) %>%
  select(id, best_lambda)

# Extract best hyperparameters for Random Forest
rf_best_params <- surgery_models %>%
  mutate(
    rf_params = map(rf, "best_params")
  ) %>%
  select(id, rf_params) %>%
  unnest(cols = c(rf_params))

# Extract best hyperparameters for GBM
gbm_best_params <- surgery_models %>%
  mutate(
    gbm_params = map(gbm, "best_params")
  ) %>%
  select(id, gbm_params) %>%
  unnest(cols = c(gbm_params))
```

## Apply the best hyper parameter to the training set

```{r final_model, echo = FALSE, warning=FALSE, message=FALSE}
# Train final models on entire training data
# Prepare data matrices for entire training data
x_train_full <- model.matrix(log(SurgeryLength)~ AgeGroupAtDateOfRemoval + CurrentPriorityCode +
                                WaitingListSpecialtyDesc + 
                                PlannedStayTypeDescription +
                                PPPCategory +
                                TreatmentCampusName + AdminCategory, data = train_data)[, -1]
y_train_full <- log(train_data$SurgeryLength)

# Ridge Regression Final Model
ridge_best_lambda_overall <- mean(ridge_best_lambdas$best_lambda)
ridge_final_model <- glmnet(
  x = x_train_full,
  y = y_train_full,
  alpha = 0,
  lambda = ridge_best_lambda_overall
)

# Lasso Regression Final Model
lasso_best_lambda_overall <- mean(lasso_best_lambdas$best_lambda)
lasso_final_model <- glmnet(
  x = x_train_full,
  y = y_train_full,
  alpha = 1,
  lambda = lasso_best_lambda_overall
)

# Random Forest Final Model
rf_best_params_overall <- rf_best_params %>%
  summarize(
    mtry = median(mtry),
    min_n = median(min_n),
    ntree = median(ntree)
  )

rf_final_model <- randomForest(
  log(SurgeryLength)~ AgeGroupAtDateOfRemoval + CurrentPriorityCode +
                                WaitingListSpecialtyDesc + 
                                PlannedStayTypeDescription +
                                PPPCategory +
                                TreatmentCampusName + AdminCategory,
  data = train_data,
  mtry = rf_best_params_overall$mtry,
  nodesize = rf_best_params_overall$min_n,
  ntree = rf_best_params_overall$ntree,
  importance = TRUE
)

# GBM Final Model
gbm_best_params_overall <- gbm_best_params %>%
  summarize(
    n.trees = median(n.trees),
    interaction.depth = median(interaction.depth),
    shrinkage = median(shrinkage),
    n.minobsinnode = median(n.minobsinnode)
  )

gbm_final_model <- gbm(
  formula = log(SurgeryLength)~ AgeGroupAtDateOfRemoval + CurrentPriorityCode +
                                WaitingListSpecialtyDesc + 
                                PlannedStayTypeDescription +
                                PPPCategory +
                                TreatmentCampusName + AdminCategory,
  data = train_data,
  distribution = "gaussian",
  n.trees = gbm_best_params_overall$n.trees,
  interaction.depth = gbm_best_params_overall$interaction.depth,
  shrinkage = gbm_best_params_overall$shrinkage,
  n.minobsinnode = gbm_best_params_overall$n.minobsinnode,
  verbose = FALSE
)
```

```{r pred_test, echo = FALSE, warning=FALSE, message=FALSE}
# Prepare test data matrices
x_test_full <- model.matrix(log(SurgeryLength) ~ AgeGroupAtDateOfRemoval + CurrentPriorityCode +
                                WaitingListSpecialtyDesc + 
                                PlannedStayTypeDescription +
                                PPPCategory +
                                TreatmentCampusName + AdminCategory, data = test_data)[, -1]
y_test_full <- test_data$SurgeryLength

# Predictions on test set
# Ridge Regression
ridge_pred_log <- predict(ridge_final_model, newx = x_test_full, s = ridge_best_lambda_overall)
ridge_pred <- exp(ridge_pred_log)

# Lasso Regression
lasso_pred_log <- predict(lasso_final_model, newx = x_test_full, s = lasso_best_lambda_overall)
lasso_pred <- exp(lasso_pred_log)

# Random Forest
rf_pred_log <- predict(rf_final_model, newdata = test_data)
rf_pred <- exp(rf_pred_log)

# GBM
gbm_pred_log <- predict(gbm_final_model, newdata = test_data, n.trees = gbm_final_model$n.trees)
gbm_pred <- exp(gbm_pred_log)

# Calculate metrics on test set
actual <- y_test_full

ridge_test_results <- data.frame(actual = actual, predicted = as.vector(ridge_pred))
lasso_test_results <- data.frame(actual = actual, predicted = as.vector(lasso_pred))
rf_test_results <- data.frame(actual = actual, predicted = rf_pred)
gbm_test_results <- data.frame(actual = actual, predicted = gbm_pred)

# Compute metrics
ridge_test_metrics <- yardstick::metrics(ridge_test_results, truth = actual, estimate = predicted) %>%
  mutate(model = "Ridge")

lasso_test_metrics <- yardstick::metrics(lasso_test_results, truth = actual, estimate = predicted) %>%
  mutate(model = "Lasso")

rf_test_metrics <- yardstick::metrics(rf_test_results, truth = actual, estimate = predicted) %>%
  mutate(model = "Random Forest")

gbm_test_metrics <- yardstick::metrics(gbm_test_results, truth = actual, estimate = predicted) %>%
  mutate(model = "GBM")

# Combine test set metrics
test_set_metrics <- bind_rows(ridge_test_metrics, lasso_test_metrics, rf_test_metrics, gbm_test_metrics)

# Pivot the metrics from long to wide format
test_set_metrics_wide <- test_set_metrics %>%
  select(model, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate)

# Display the wide format table
print(test_set_metrics_wide)
```