---
title: "Presentation_Plot"
author: "Phuong Anh Mai"
date: "2024-10-29"
output: html_document
---

This R markdown files contain all plot and table code for the Patients Finder Tool Presentation. 

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(readr)
library(dplyr)
library(tidymodels)
library(stringdist)
library(glmnet)
library(randomForest)
library(gbm)
library(yardstick)
library(ggplot2)
library(plotly)
library(gridExtra)
library(kableExtra)
library(knitr)
library(visdat)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
```

```{r reading datasets, echo = FALSE, warning=FALSE, message = FALSE}
waiting_list <- read_csv("data/new_waiting_list.csv")
surgery_patients <- read.csv("data/joined_data.csv")
```

# Data wrangling 
## Convert Data types and Filter based on assumptions

```{r wrangling, echo = FALSE, warning=FALSE, message=FALSE}
# Filter out patient > 17 years old
surgery_patients <- surgery_patients %>% 
  filter(AgeAtDateOfRemoval >17)

# Getting waiting list data
waiting_patients <- waiting_list %>% 
  filter(ReportedRemovalDateTime == "NULL",
         RemovalDateTime == "NULL")

# Remove or impute missing data 
surgery_patients <- na.omit(surgery_patients)

# Remove unwanted columns
unwanted <- c("Paed ENT",
"Paed General Surgery",
"Paed Oral & Maxillofacial Surgery",
"Paed Orthopaedic Surgery",
"Paed Plastic Surgery",
"Paed Urology")

surgery_patients <- surgery_patients[!surgery_patients$WaitingListSpecialtyDesc %in% unwanted, ]

# Identify duplicates
duplicates <- duplicated(surgery_patients)
# Remove duplicates
surgery_patients <- surgery_patients[!duplicated(surgery_patients), ]

# glimpse(surgery_patients)

# Create PPPCategory for both datasets
surgery_patients <- surgery_patients %>%
  mutate(PPPCategory = case_when(
    grepl("arthro|joint|shoulder|knee|ligament|rotator cuff|ACL|hip|meniscus|spine|fusion|discectomy", 
          PPPDesc, ignore.case = TRUE) ~ "Orthopedic",
    grepl("heart|cardiac|bypass|coronary|valve|CABG|aortic|angioplasty|stent|aneurysm", 
          PPPDesc, ignore.case = TRUE) ~ "Cardiac",
    grepl("appendic|cholecyst|bowel|hernia|gastrectomy|colectomy|colorectal|laparotomy|gastrostomy|fundoplication", 
          PPPDesc, ignore.case = TRUE) ~ "General Surgery",
    grepl("breast|mastectomy|lumpectomy|reconstruction|augmentation|reduction|prosthesis|biopsy|oncology", 
          PPPDesc, ignore.case = TRUE) ~ "Breast Surgery",
    grepl("urology|kidney|nephrectomy|bladder|prostate|ureter|urethra|renal", 
          PPPDesc, ignore.case = TRUE) ~ "Urology",
    grepl("ENT|sinus|tonsil|adenoid|larynx|otoplasty|nose|nasal|septoplasty|ear|throat", 
          PPPDesc, ignore.case = TRUE) ~ "ENT Surgery",
    grepl("gynaecology|hysterectomy|cervix|uterus|fallopian|ovary|endometrial|ablation|vaginal", 
          PPPDesc, ignore.case = TRUE) ~ "Gynaecology",
    grepl("neurosurgery|brain|craniotomy|spinal|meninges|nerve|tumor|neuro", 
          PPPDesc, ignore.case = TRUE) ~ "Neurosurgery",
    TRUE ~ "Other"  # Default category if none of the above matches
  ))


waiting_list <- waiting_list %>%
  mutate(PPPCategory = case_when(
    grepl("arthro|joint|shoulder|knee|ligament|rotator cuff|ACL|hip|meniscus|spine|fusion|discectomy", 
          PPPDesc, ignore.case = TRUE) ~ "Orthopedic",
    grepl("heart|cardiac|bypass|coronary|valve|CABG|aortic|angioplasty|stent|aneurysm", 
          PPPDesc, ignore.case = TRUE) ~ "Cardiac",
    grepl("appendic|cholecyst|bowel|hernia|gastrectomy|colectomy|colorectal|laparotomy|gastrostomy|fundoplication", 
          PPPDesc, ignore.case = TRUE) ~ "General Surgery",
    grepl("breast|mastectomy|lumpectomy|reconstruction|augmentation|reduction|prosthesis|biopsy|oncology", 
          PPPDesc, ignore.case = TRUE) ~ "Breast Surgery",
    grepl("urology|kidney|nephrectomy|bladder|prostate|ureter|urethra|renal", 
          PPPDesc, ignore.case = TRUE) ~ "Urology",
    grepl("ENT|sinus|tonsil|adenoid|larynx|otoplasty|nose|nasal|septoplasty|ear|throat", 
          PPPDesc, ignore.case = TRUE) ~ "ENT Surgery",
    grepl("gynaecology|hysterectomy|cervix|uterus|fallopian|ovary|endometrial|ablation|vaginal", 
          PPPDesc, ignore.case = TRUE) ~ "Gynaecology",
    grepl("neurosurgery|brain|craniotomy|spinal|meninges|nerve|tumor|neuro", 
          PPPDesc, ignore.case = TRUE) ~ "Neurosurgery",
    TRUE ~ "Other"  # Default category 
  ))

# Convert data types
surgery_patients <- surgery_patients %>%
  mutate(
    PlannedTheatreTimeInMins = as.numeric(PlannedTheatreTimeInMins),
    ArrivedTheatreDateTime = as.POSIXct(ArrivedTheatreDateTime, format = "%Y-%m-%d %H:%M:%S"),
    OutOfTheatreDateTime = as.POSIXct(OutOfTheatreDateTime, format = "%Y-%m-%d %H:%M:%S"),
    WaitingStartDateTime = as.POSIXct(WaitingStartDateTime, format = "%Y-%m-%d %H:%M:%S"),
    SurgeryStartDateTime = as.POSIXct(SurgeryStartDateTime, format = "%Y-%m-%d %H:%M:%S"),
    SurgeryEndDateTime = as.POSIXct(SurgeryEndDateTime, format = "%Y-%m-%d %H:%M:%S"),
    IntoTheatreDateTime = as.POSIXct(IntoTheatreDateTime, format = "%Y-%m-%d %H:%M:%S"),
    AgeGroupAtDateOfRemoval = as.factor(AgeGroupAtDateOfRemoval),
    CurrentPriorityCode = as.factor(CurrentPriorityCode),
    PPPCategory = as.factor(PPPCategory),
    # Use for model fitting
    WaitingListSpecialtyDesc = as.factor(WaitingListSpecialtyDesc),
    OperatingDrCarerName = as.factor(OperatingDrCarerName),
    TreatmentCampusName = as.factor(TreatmentCampusName),
    PlannedStayTypeDescription = as.factor(PlannedStayTypeDescription),
    AdminCategory = as.factor(AdminCategory),
  )

# Calculate real-time of surgery
surgery_patients <- surgery_patients %>% 
  mutate(SurgeryLength = round((difftime(OutOfTheatreDateTime, IntoTheatreDateTime, units = "mins")),4)) %>% 
  mutate(SurgeryLength = as.numeric((SurgeryLength)))

# Convert Surgery Length data types
surgery_patients <- surgery_patients %>%
mutate(SurgeryLength = as.numeric(SurgeryLength))

waiting_list <- waiting_list %>%
  mutate(
    AgeGroupAtDateOfRemoval = as.factor(AgeGroupAtDateOfRemoval),
    CurrentPriorityCode = as.factor(CurrentPriorityCode),
    WaitingListSpecialtyDesc = as.factor(WaitingListSpecialtyDesc),
    PPPCategory = as.factor(PPPCategory),
    TreatmentCampusName = as.factor(TreatmentCampusName),
    PlannedStayTypeDescription = as.factor(PlannedStayTypeDescription),
    AdminCategory = as.factor(AdminCategory)
  )
```

## Further refine of the dataset

```{r surgery length, echo = FALSE, warning=FALSE, message=FALSE}}
# Remove Specialty with under 10 cases
unwanted_specialty<- surgery_patients %>% 
  group_by(WaitingListSpecialtyDesc) %>%
  summarise(CaseCount = n()) %>% 
  filter(CaseCount < 10)

surgery_patients <- surgery_patients[!surgery_patients$WaitingListSpecialtyDesc %in% unwanted_specialty$WaitingListSpecialtyDesc, ]

# Remove negative values and outliers
# Filter out Surgery within 10 hours, as the longer surgery requested more complex estimation
surgery_patients <- surgery_patients %>% 
  na.omit() %>% 
  filter(SurgeryLength > 10) %>% 
  filter(SurgeryLength < 600) %>% 
  filter(PlannedTheatreTimeInMins > 10) %>% 
  filter(PlannedTheatreTimeInMins < 600)
```

# Exploratory Data Analysis (EDA)
## Observe the human estimation and real-time of Surgery Length

```{r distribution visualize, echo = FALSE, warning=FALSE, message=FALSE}
# Visualize the distribution of Planned Theatre Time
a <- ggplot(surgery_patients, aes(x = PlannedTheatreTimeInMins)) +
  geom_histogram(fill = "blue", color = "white", bins = 30) +
  theme_minimal() +
  labs(title = "Distribution of Planned Theatre Time in Minutes",
       x = "Planned Theatre Time (mins)")

# Visualize the distribution of Time Difference
b <- ggplot(surgery_patients, aes(x = SurgeryLength)) +
  geom_histogram( fill = "#E69F00", color = "white", bins = 30) +
  theme_minimal() +
  labs(title = "Distribution of Surgery Time in Minutes",
       x = "Operating Theatre Time (mins)")

gridExtra::grid.arrange(a, b)
```

## Investigate the Surgery Length by Specialty

```{r visualize boxplot, echo = FALSE, warning=FALSE, message=FALSE}
# Boxplot of SurgeryLength with an additional geom_line for average
ggplot(surgery_patients, aes(x = SurgeryLength, y = fct_reorder(WaitingListSpecialtyDesc, SurgeryLength, .fun = median))) +
  geom_boxplot(fill = "#E69F00") +
  
  theme_minimal() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 55, hjust = 1)) +
  labs(title = "Surgery Length by Specialty",
       x = "Surgery Length (mins)",
       y = "Specialty")
```

## Observe the Waiting Days

```{r Waiting days, echo = FALSE, warning=FALSE, message=FALSE}
# Calculating Waiting day variables 
surgery_patients <- surgery_patients %>%
  mutate(WaitingDays = as.numeric(round(difftime(SurgeryStartDateTime, WaitingStartDateTime, units = "days"))))

# Proportion of people in waiting within 1 year
proportion_within_1_year <- surgery_patients %>%
  summarise(Proportion = round(mean(WaitingDays <= 365),4))

# Average waiting time for particular Specialty
Average_waiting_day_by_specialty <- surgery_patients %>%
  group_by(WaitingListSpecialtyDesc) %>%
  summarise(Avearge_waiting_day = round(mean(WaitingDays)),
            CaseCount = n()) %>% 
  arrange(desc(Avearge_waiting_day))

# Plot Average waiting time for particular Specialty
p <- ggplot(Average_waiting_day_by_specialty, aes(x = reorder(WaitingListSpecialtyDesc, Avearge_waiting_day), y = Avearge_waiting_day)) +
  geom_segment(aes(xend = reorder(WaitingListSpecialtyDesc, Avearge_waiting_day), y = 0, yend = Avearge_waiting_day), color = "grey") +
  geom_point(aes(size = CaseCount), color = "#E69F00")  +
  theme_minimal() +
  labs(title = "Average Waiting Day by Specialty",
       x = "Specialty",
       y = "Avearge waiting day") +
  theme(axis.text.x = element_text(angle = 55, hjust = 1))

plotly_plot <- ggplotly(p)

plotly_plot
```

# Feature engineering: Extract interesting features
## Calculate and plot the Average Surgery length by Age Group

```{r}
# Calculating Average Surgery Length by Age Group
AverageSurgeryLength_age <- surgery_patients %>% 
  group_by(AgeGroupAtDateOfRemoval) %>% 
  summarise(AverageSurgeryLength_age = round(mean(SurgeryLength, na.rm = TRUE))) %>% 
  arrange(desc(AverageSurgeryLength_age))

# Displaying the result
ggplot(AverageSurgeryLength_age, aes(x = AgeGroupAtDateOfRemoval, y = AverageSurgeryLength_age)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength_age), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Age Group", 
       x = "Age Groups", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10)
  )
```

## Calculate and plot the Average Surgery length by Specialty

```{r}
# Calculate the average of SurgeryLength by specialty
Average_surgery_length_by_specialty <- surgery_patients %>%
  group_by(WaitingListSpecialtyDesc) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the result
ggplot(Average_surgery_length_by_specialty, aes(x = reorder(WaitingListSpecialtyDesc, -AverageSurgeryLength), y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Specialty", 
       x = "Specialty", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10)
  )
```

### Line graph for Average Surgery Length by Specialty
```{r echo = FALSE, warning=FALSE, message=FALSE}
# Investigate relationships between Specialty and SurgeryLength
ggplot() +
  # Add geom_line using the average surgery length by specialty
  geom_line(data = Average_surgery_length_by_specialty, aes(x = AverageSurgeryLength, 
                                               y = fct_reorder(WaitingListSpecialtyDesc, AverageSurgeryLength),
                                               group = 1), color = "blue", size = 1) +
   theme_minimal() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 55, hjust = 1)) +
  labs(title = "Average Surgery Length by Specialty",
       x = "Average Surgery Length (mins)",
       y = "Specialty")
```

## Calculate and plot the Average Surgery length by Admin category

```{r}
# Calculate average surgery length by Admin category
Average_surgery_length_admin <- surgery_patients %>%
  group_by(AdminCategory) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the result
ggplot(Average_surgery_length_admin, aes(x = reorder(AdminCategory, -AverageSurgeryLength), y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Admin Category", 
       x = "Admin Category", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10) 
  )
```

## Calculate and plot the Average Surgery length by Planned Stay

```{r}
# Calculate average surgery length by Planned Stay
Average_surgery_length_plan_stay <- surgery_patients %>%
  group_by(PlannedStayTypeDescription) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Displaying the result
Average_surgery_length_plan_stay %>%
  kable("html", caption = "Average Surgery Length by Planned Stay") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = FALSE, 
                font_size = 14) %>%
  column_spec(1, bold = TRUE, border_right = TRUE) %>%  
  column_spec(2, color = "blue", background = "lightgrey") %>% 
  add_header_above(c(" " = 1, " " = 1)) 
```

## Calculate and plot the Average Surgery length by Treatment Campus

```{r}
# Calculate average surgery length by Treatment Campus
Average_surgery_length_treatment_campus <- surgery_patients %>%
  group_by(TreatmentCampusName) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the results
ggplot(Average_surgery_length_treatment_campus, aes(x = reorder(TreatmentCampusName, -AverageSurgeryLength), y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Treatment Campus", 
       x = "Treatment Campus", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10)
  )
```

## Calculate and plot the Average Surgery length by Current Priority Codes

```{r}
# Calculate average surgery length by Current Priority Codes
Average_surgery_length_priority <- surgery_patients %>%
  group_by(CurrentPriorityCode) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the results
ggplot(Average_surgery_length_priority, aes(x = CurrentPriorityCode, y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +
  coord_flip() +
  labs(title = "Average Surgery Length by Priority Code", 
       x = "Priority Code", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.margin = ggplot2::margin(10, 10, 10, 10)
  )
```

## Calculate and plot the Average Surgery Length by PPP Category

```{r PPP Desc by group}
# Calculate average surgery length by PPP Category
Average_surgery_length_ppp_category <- surgery_patients %>%
  group_by(PPPCategory) %>%
  summarise(AverageSurgeryLength = round(mean(SurgeryLength, na.rm = TRUE)))%>% 
  arrange(desc(AverageSurgeryLength))

# Visualizing the result
ggplot(Average_surgery_length_ppp_category, aes(x = reorder(PPPCategory, -AverageSurgeryLength), y = AverageSurgeryLength)) +
  geom_bar(stat = "identity", fill = "#E69F00", width = 0.7) +  
  geom_text(aes(label = AverageSurgeryLength), hjust = -0.2, color = "black", size = 3.5) +  
  coord_flip() +  
  labs(title = "Average Surgery Length by PPP Category", 
       x = "PPP Category", 
       y = "Average Surgery Length (minutes)") +
  theme_minimal() +  # Clean theme
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  
    axis.text.y = element_text(size = 12),  
    axis.title.y = element_text(size = 14),  
    axis.title.x = element_text(size = 14),  
    plot.margin = ggplot2::margin(10, 10, 10, 10)  
  )
```